---
title: "DATA 621 – Business Analytics and Data Mining"
author: "Abdelmalek Hajjam/ Monu Chacko"
date: "4/26/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

    
#### In this homework assignment, you will explore, analyze and model a dataset containing approximately 8000 records representing  a  customer  at  an  auto  insurance  company. Each  record  has  two response variables.  The first responsevariable, TARGET_FLAG, is a 1 or a 0. A “1” means that the person was in a car crash. A zero means that the person was not in a car crash.The second responsevariable is TARGET_AMT. This value is zero if the person did not crash their car. But if they did crash their car, this number will be a value greater than zero.

    
#### Your objective is to build multiple linear regression and binary logistic regression models on the training data to predict the probability that a person will crash their car and also the amount of money it will cost if the person does crash their car. You can only use the variables given to you (or variables that you derive from the variables provided). Below is a short description of the variables of interest in the data set:

     


```{r message=FALSE}
library(ggcorrplot)
library(car)
library(MASS)
library(dplyr)
library(ggplot2)
library(caret)
library(pROC)
library(pscl)
library(psych)
library(data.table) 
```

### DATA EXPLORATION

```{r}
# Read data
train_df <- read.csv("https://raw.githubusercontent.com/monuchacko/cuny_msds/master/data_621/Homework4/insurance_training_data.csv", stringsAsFactors = FALSE)

dim(train_df)
str(train_df)
head(train_df)
```


```{r}
# Exclude the INDEX column
tr <- train_df[-1]

#  Convert to numeric
tr$INCOME <- as.numeric(gsub('[$,]', '', tr$INCOME))
tr$HOME_VAL <- as.numeric(gsub('[$,]', '', tr$HOME_VAL))
tr$BLUEBOOK <- as.numeric(gsub('[$,]', '', tr$BLUEBOOK))
tr$OLDCLAIM <- as.numeric(gsub('[$,]', '', tr$OLDCLAIM))

# Remove characters that are not required
tr$MSTATUS  <- gsub("z_", "", tr$MSTATUS)
tr$SEX  <- gsub("z_", "", tr$SEX)
tr$EDUCATION  <- gsub("z_", "", tr$EDUCATION)
tr$JOB  <- gsub("z_", "", tr$JOB)
tr$CAR_USE  <- gsub("z_", "", tr$CAR_USE)
tr$CAR_TYPE  <- gsub("z_", "", tr$CAR_TYPE)
tr$URBANICITY  <- gsub("z_", "", tr$URBANICITY)


# Reorder columns -- predictor categorical, predictor numeric, target
indx <- c(8, 10:13, 15, 18:19, 22, 25, 3:7, 9, 14, 16:17, 20:21, 23:24, 1:2)
tr_ordered <- tr
setcolorder(tr_ordered,indx)
```


#### Examine the data

```{r}
table(tr$PARENT1)
table(tr$MSTATUS)
table(tr$SEX)
table(tr$EDUCATION)
table(tr$JOB)
table(tr$CAR_USE)
table(tr$CAR_TYPE)
table(tr$RED_CAR)
table(tr$REVOKED)
table(tr$URBANICITY)
```

#### Summary of the data

```{r}
summary(tr)
```


```{r}
boxplot((INCOME/1000)~AGE,data=tr, main="Income vs Age", xlab="Age", ylab="Income")
```


### DATA PREPARATION

```{r}
tr_prep <- tr_ordered
```


```{r}
M <- sapply(tr_prep, function(x) sum(x=="") | sum(is.na(x))); names(M[(M>0)])
```

```{r}
x <- c(12, 14, 15, 16, 23)
par(mfrow=c(2,3))
for (val in x) {
  hist(tr_prep[,val],xlab=names(tr_prep[val]), main="")
}
par(mfrow=c(1,1))
```



```{r}
#impute

tr_prep = tr_prep %>% 
  mutate(AGE = 
           ifelse(is.na(AGE), 
                  mean(AGE, na.rm=TRUE), AGE)) %>% 

  mutate(YOJ = 
           ifelse(is.na(YOJ), 
                  mean(YOJ, na.rm=TRUE), YOJ)) %>% 

  mutate(INCOME = 
           ifelse(is.na(INCOME), 
                  median(INCOME, na.rm=TRUE), INCOME)) %>% 

  mutate(HOME_VAL = 
           ifelse(is.na(HOME_VAL), 
                  mean(HOME_VAL, na.rm=TRUE), HOME_VAL)) %>% 

  mutate(CAR_AGE = 
           ifelse(is.na(CAR_AGE), 
                  mean(CAR_AGE, na.rm=TRUE), CAR_AGE)) %>% 

  mutate(JOB = 
           ifelse((JOB == "" & EDUCATION == 'PhD'),
                  "Doctor", JOB)) %>% 

  mutate(JOB = 
           ifelse((JOB == "" & EDUCATION == 'Masters'),
                  "Lawyer", JOB))
```

```{r}
M <- sapply(tr_prep, function(x) sum(x=="") | sum(is.na(x))); names(M[(M>0)])
```


```{r}
# Outlier Capping

tr_prep2 <- tr_prep

id <- c(11:23)
for (val in id) {
  qnt <- quantile(tr_prep2[,val], probs=c(.25, .75), na.rm = T)
  caps <- quantile(tr_prep2[,val], probs=c(.05, .95), na.rm = T)
  H <- 1.5 * IQR(tr_prep2[,val], na.rm = T)
  tr_prep2[,val][tr_prep2[,val] < (qnt[1] - H)] <- caps[1]
  tr_prep2[,val][tr_prep2[,val] > (qnt[2] + H)] <- caps[2]

}
```


## BUILD MODELS

```{r}
nTrain <- createDataPartition(tr_prep2$TARGET_FLAG, p=0.8, list=FALSE)
ntraining <- tr_prep2[nTrain,]
ntesting <- tr_prep2[-nTrain,]

set.seed(123)
```



```{r}
# Logistic Regression build the model using training set
full.model_FLAG <- glm(TARGET_FLAG ~.-TARGET_AMT, data = ntraining , family = binomial)
summary(full.model_FLAG)
```



```{r}
round(exp(cbind(Estimate=coef(full.model_FLAG))),2)
```


```{r}
# evaluate the model by predicting using the testing set
m1_prob <- predict(full.model_FLAG, ntesting, type = "response")
m1_pclass <- ifelse(m1_prob >= 0.5, 1, 0)

# create confusion matrix
pclass <- factor(m1_pclass,levels = c(1,0))
aclass <- factor(ntesting$TARGET_FLAG,levels = c(1,0))
confusionMatrix(pclass, aclass);
```


```{r}
# plot and show area under the curve
plot(roc(ntesting$TARGET_FLAG, m1_prob),print.auc=TRUE)
```


```{r}
# get McFadden
m1_mcFadden <- pR2(full.model_FLAG); m1_mcFadden["McFadden"]
```



```{r}
full.model.AMT <- lm(TARGET_AMT ~. -TARGET_FLAG,  data = tr_prep2)
summary(full.model.AMT)
```



```{r}
vif(full.model.AMT)
```


```{r}
par(mfrow=c(2,2))
plot(full.model.AMT)
```


```{r}
par(mfrow=c(1,1))
```


### 3.2 Stepwise variable selection


```{r}
# Logistic Regression build the model using training set
step.model_FLAG <- full.model_FLAG %>% stepAIC(trace = FALSE)
summary(step.model_FLAG)
```


```{r}
round(exp(cbind(Estimate=coef(step.model_FLAG))),2)
```


```{r}
# evaluate the model by predicting using the testing set
m2_prob <- predict(step.model_FLAG, ntesting, type = "response")
m2_pclass <- ifelse(m2_prob >= 0.5, 1, 0)

# create confusion matrix
pclass <- factor(m2_pclass,levels = c(1,0))
aclass <- factor(ntesting$TARGET_FLAG,levels = c(1,0))
confusionMatrix(pclass, aclass);

```




```{r}
# plot and show area under the curve
plot(roc(ntesting$TARGET_FLAG, m2_prob),print.auc=TRUE)
```


```{r}
# get McFadden
m2_mcFadden <- pR2(step.model_FLAG); m2_mcFadden["McFadden"]
```



```{r}
# Linear Regression - TARGET_AMT
step.model.AMT <- full.model.AMT %>% stepAIC(trace = FALSE)
summary(step.model.AMT)
```


```{r}
vif(step.model.AMT)
```


```{r}
par(mfrow=c(2,2))
plot(step.model.AMT)
```


```{r}
par(mfrow=c(1,1))
```


### 3.3 Significant value variable selection

```{r}
# Logistic Regression build the model using training set
select.model_FLAG <- glm(TARGET_FLAG ~.
                       -TARGET_AMT
                       -EDUCATION
                       -SEX
                       -RED_CAR
                       -KIDSDRIV
                       -AGE
                       -HOMEKIDS
                       -YOJ
                       -HOME_VAL
                       -OLDCLAIM
                       -BLUEBOOK 

                         , data = ntraining , family = binomial)
summary(select.model_FLAG)
```


```{r}
round(exp(cbind(Estimate=coef(select.model_FLAG))),2)
```


```{r}
# evaluate the model by predicting using the testing set
m3_prob <- predict(select.model_FLAG, ntesting, type = "response")
m3_pclass <- ifelse(m3_prob >= 0.5, 1, 0)

# create confusion matrix
pclass <- factor(m3_pclass,levels = c(1,0))
aclass <- factor(ntesting$TARGET_FLAG,levels = c(1,0))
confusionMatrix(pclass, aclass);
```


```{r}
# plot and show area under the curve
plot(roc(ntesting$TARGET_FLAG, m3_prob),print.auc=TRUE)
```


```{r}
# get McFadden
m3_mcFadden <- pR2(select.model_FLAG); m3_mcFadden["McFadden"]
```


```{r}
# Linear Regression - TARGET_AMT
select.model.AMT <- lm(TARGET_AMT ~. 
                       -TARGET_FLAG
                       -EDUCATION
                       -SEX
                       -RED_CAR
                       -KIDSDRIV
                       -AGE
                       -HOMEKIDS
                       -YOJ
                       -HOME_VAL
                       -OLDCLAIM
                       -BLUEBOOK 
                       ,  data = tr_prep2)
summary(select.model.AMT)
```

```{r}
vif(select.model.AMT)
```


```{r}
par(mfrow=c(2,2))
plot(select.model.AMT)
```


```{r}
par(mfrow=c(1,1))
```


### SELECT MODELS

```{r}

# Read the evaluation dataset
 eval_df <- read.csv("https://raw.githubusercontent.com/monuchacko/cuny_msds/master/data_621/Homework4/insurance-evaluation-data.csv", stringsAsFactors = FALSE)


# Remove columns not selected in 2nd model
#eval_df <- dplyr::select(eval_df, -YOJ, -MSTATUS, -RED_CAR)

#  Convert to numeric
eval_df$INCOME <- as.numeric(gsub('[$,]', '', eval_df$INCOME))
eval_df$HOME_VAL <- as.numeric(gsub('[$,]', '', eval_df$HOME_VAL))
eval_df$BLUEBOOK <- as.numeric(gsub('[$,]', '', eval_df$BLUEBOOK))
eval_df$OLDCLAIM <- as.numeric(gsub('[$,]', '', eval_df$OLDCLAIM))

# Remove irrelevant characters
eval_df$MSTATUS  <- gsub("z_", "", eval_df$MSTATUS)
eval_df$SEX  <- gsub("z_", "", eval_df$SEX)
eval_df$EDUCATION  <- gsub("z_", "", eval_df$EDUCATION)
eval_df$JOB  <- gsub("z_", "", eval_df$JOB)
eval_df$CAR_USE  <- gsub("z_", "", eval_df$CAR_USE)
eval_df$CAR_TYPE  <- gsub("z_", "", eval_df$CAR_TYPE)
eval_df$URBANICITY  <- gsub("z_", "", eval_df$URBANICITY)

#impute
eval_df = eval_df %>% 
  mutate(AGE = 
           ifelse(is.na(AGE), 
                  mean(AGE, na.rm=TRUE), AGE)) %>% 

  mutate(YOJ = 
           ifelse(is.na(YOJ), 
                  mean(YOJ, na.rm=TRUE), YOJ)) %>% 

  mutate(INCOME = 
           ifelse(is.na(INCOME), 
                  median(INCOME, na.rm=TRUE), INCOME)) %>% 

  mutate(HOME_VAL = 
           ifelse(is.na(HOME_VAL), 
                  mean(HOME_VAL, na.rm=TRUE), HOME_VAL)) %>% 

  mutate(CAR_AGE = 
           ifelse(is.na(CAR_AGE), 
                  mean(CAR_AGE, na.rm=TRUE), CAR_AGE)) %>% 

  mutate(JOB = 
           ifelse((JOB == "" & EDUCATION == 'PhD'),
                  "Doctor", JOB)) %>% 

  mutate(JOB = 
           ifelse((JOB == "" & EDUCATION == 'Masters'),
                  "Lawyer", JOB))


 eval_prob <- predict(step.model_FLAG, eval_df, type = "response")
 eval_pclass <- ifelse(eval_prob >= 0.5, 1, 0)
 
 eval_amt <- ifelse(eval_pclass == 1, predict(step.model.AMT, eval_df, type = "response"), 0)
 
 
 eval_df$TARGET_FLAG <- eval_pclass
 eval_df$TARGET_AMT <- eval_amt
  
 head(eval_df)
 
# Export
#  write.csv(eval_df,file="Insurance_Results.csv")
  
```



